---
title: "p8105_hw1_js5095"
author: Jianting Shi
date: 09-17-2020
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1

Create a data frame with the specified elements.

```{r create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    logi_vec = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    fact_vec = factor(c("up", "up", "up", "down", "down", "down", "down", "same", "same", "same")),
  )
```

Take the mean of each variable in the data frame.

```{r mean_pull_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, logi_vec))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, fact_vec))
```

The function mean() only works in numeric and logical variables but not character or factor.

```{r numeric_pull_var, results='hide', warning=FALSE}
as.numeric(pull(prob1_df, logi_vec))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, fact_vec))
```

The function as.numeric() can be applied to logical and factor variables but not character, as logical and factor variables contain indexes that can be converted to numeric values. However, factor variable with character levels need to be converted to numeric levels first. That may be the reason why mean() could only work on numeric and logical variables.

```{r numeric_pull_var_2}
as.numeric(pull(prob1_df, logi_vec)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, logi_vec)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, logi_vec))) * pull(prob1_df, samp)
```

## Problem 2

```{r}
data("penguins", package = "palmerpenguins")
```

The dataset "penguins" has multiple variables including `r ls(penguins)`. \
Here are a preview of the dataset: \
`r capture.output(str(penguins))`. \
The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. \
The mean of flipper length is `r mean(na.omit(pull(penguins, flipper_length_mm)))` (excluding NA from the data).

```{r plot, echo=FALSE, warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

